---
title: "Agribusiness Management and Cooperative"
subtitle: "Practical"
author: "Deependra Dhakal"
date: "11/30/2020"
output: 
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 1
    keep_tex: yes
    includes:
      in_header: header_template.tex
---

```{r setup, include=FALSE}
require(tidyverse)
library(knitr)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "55%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2, knitr.kable.NAN = "")
options(scipen = 40)
# options(kableExtra.latex.load_packages = FALSE)

```

\clearpage

# Determination of optimum input use and maximization of profit using one input

## Objectives {-}

- To determine optimum input use and maximize profit using one input

**Numerical**

1. For the following hypothetical data showing input-output relationship between number of tractor plus driver units recruits and hectares of land ploughed, determine.

```{r tractor-driver-unit-cmr}
tractor_driver_unit_cmr <- tribble(~"tractor_driver_unit", ~"field_ploughed", 
                                   1, 2, 
                                   2, 4,
                                   5, 10, 
                                   6, 12) #%>% 
  # mutate(marginal_tractor_driver_unit = tsibble::difference(tractor_driver_unit),
  #        marginal_field_ploughed = tsibble::difference(field_ploughed), 
  #        marginal_rate_returns = field_ploughed/tractor_driver_unit)
```


```{r tractor-driver-unit-cmr-tab}
tractor_driver_unit_cmr %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  # knitr::kable(booktab = TRUE, caption = "Constant marginal rate of returns for a input-output relationship between number of tractor plus driver unit recruits and hectares of land ploughed.", 
  #              align = "c") %>% 
  knitr::kable(booktab = TRUE, caption = "Input-output relationship between number of tractor plus driver unit recruits and hectares of land ploughed.", 
               align = "c") %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position")
```

a. Which is input ?
b. Which is output ?
c. Marginal rate of returns at each level of input
d. What is the relationship between input-output ?

2. For the following hypothetical data showing input-output relationship between wheat seed used per hectare and wheat production, determine.

```{r nitrogen-wheat-imr}
nitrogen_wheat_imr <- tribble(~"wheat_seed", ~"marginal_wheat_seed", ~"wheat_production", ~"marginal_wheat_production",
        10, NA, 1000, NA,
        15, NA, 1025, NA,
        20, NA, 1075, NA,
        25, NA, 1150, NA,
        30, NA, 1250, NA,
        35, NA, 1375, NA) #%>% 
  # mutate(marginal_wheat_seed = tsibble::difference(wheat_seed), 
  #        marginal_wheat_production = tsibble::difference(wheat_production), 
  #        marginal_rate_returns = marginal_wheat_production/marginal_wheat_seed)
```


```{r nitrogen-wheat-imr-tab}
nitrogen_wheat_imr %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  # knitr::kable(booktabs = TRUE, caption = "Increasing marginal rate of return for hypothetical wheat production scenario.") %>%
  knitr::kable(booktabs = TRUE, caption = "Marginal rate of return for hypothetical wheat production scenario.") %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position")

```

a. Which is input ?
b. Which is output ?
c. Marginal rate of returns at each level of input
d. Trend of marginal rate of return

3. For the following hypothetical data showing input-output relationship between fertilizer use in wheat per hectare and wheat production, determine.

```{r nitrogen-wheat-dmr}
nitrogen_wheat_dmr <- tribble(~"wheat_fertilizer", ~"marginal_wheat_fertilizer", ~"wheat_production", ~"marginal_wheat_production",
        100, NA, 2000, NA,
        150, NA, 2600, NA,
        200, NA, 3100, NA,
        250, NA, 3500, NA,
        300, NA, 3800, NA,
        350, NA, 4000, NA) #%>% 
  # mutate(marginal_wheat_fertilizer = tsibble::difference(wheat_fertilizer), 
  #        marginal_wheat_production = tsibble::difference(wheat_production), 
  #        marginal_rate_returns = marginal_wheat_production/marginal_wheat_fertilizer)
```


```{r nitrogen-wheat-dmr-tab}
nitrogen_wheat_dmr %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  # knitr::kable(booktabs = TRUE, caption = "Decreasing marginal rate of return for hypothetical wheat production scenario.") %>% 
  knitr::kable(booktabs = TRUE, caption = "Marginal rate of return for hypothetical wheat production scenario.") %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position")

```

a. Which is input ?
b. Which is output ?
c. Marginal rate of returns at each level of input
d. Trend of marginal rate of return

\clearpage

# Determination of least cost combination of inputs using graphical and arithmetic method

## Objectives {-}

- To graphically represent cost combinations for two inputs
- To arithmetically determine the least cost combination of inputs

**Numerical**

1. For the following hypothetical data showing relationship between two fertilizers (Nitrogen and Phosphorus) used in Wheat crop to produce common yield of 2 tons per hectare, determine:

```{r}
tribble(
  ~"yield", ~"nitrogen", ~"phosphorus", 
  "2 tons", 46, 0,
  "2 tons", 32, 2,
  "2 tons", 20, 4,
  "2 tons", 10, 6,
  "2 tons", 1, 8,
  "2 tons", 0, 10
) %>% 
  mutate(delta_n_x2 = tsibble::difference(nitrogen, 1, 1), 
         delta_n_x1 = tsibble::difference(phosphorus, 1, 1), 
         mrs_x2_x1 = abs(delta_n_x2/delta_n_x1)) %>% 
  knitr::kable(booktabs = TRUE, 
               col.names = c("Yield", "Nitrogen ($X_2$)", "Phosphorus ($X_1$)", 
                             "$\\Delta X_2$", 
                             "$\\Delta X_1$", 
                             "MRS ($\\frac{\\Delta X_2}{\\Delta X_1}$)"
                             ), 
               escape = FALSE) %>% 
  kableExtra::kable_styling(latex_options = c("striped", "HOLD_position"), font_size = 8)
```


a. Marginal rate of input substitution of nitrogen for phosphorus
b. Graph showing cost combination of inputs and the least cost combination point.
c. Calculate the least cost combination of input use using arithmetic method if the price of Phosphorus is Rs 40/kg and price of Nitrogen is Rs 20/kg.

\clearpage

# Determination of optimum enterprise combination for revenue maximization

## Objectives {-}

- To maximize revenue through optimum enterprise combination

**Numerical**

For the given hypothetical data on alternative combinations of Rice and Wheat enterprise using same amount of resources, determine:

```{r pp-schedule}
pps_less <- readxl::read_xlsx("./../data/concepts_of_economics.xlsx", 
                               sheet = "production_possibility_schedule", col_names = TRUE)
pps_less %>% 
  mutate_if(is.numeric, function(x) {
    kableExtra::cell_spec(x, "latex", bold = T, color = kableExtra::spec_color(x, end = 0.8),
                          font_size = kableExtra::spec_font_size(x))
    }) %>%
  mutate_at("Alternative outputs", function(x) {kableExtra::cell_spec(
    x, "latex", color = "white", bold = T,
    background = kableExtra::spec_color(seq_along(1:nrow(.)), end = 0.9, option = "A", direction = -1)
    )}) %>% 
  knitr::kable(format = "latex", caption = "Production possibility schedule", 
               escape = F, booktabs = T, linesep = "", align = "c") %>% 
  kableExtra::kable_styling(position = "c", full_width = FALSE, font_size = 8, latex_options = "HOLD_position") %>% 
  kableExtra::column_spec(1:3, width = "10em")
```


1. Graphically show production possibility frontier.
2. Which type of product combination is indicated ?
3. Using arithmetic method determine optimum production with combination of enterprises. Assume that price of Rice is Rs 20/kg and price of Wheat is Rs 30/kg.

\clearpage

# Farm record keeping and preparation of farm inventory

## Objectives {-}

- To understand usefulness and purpose of keeping farm record
- To be able to maintain farm inventory of an operating farm

## Introduction {-}

Farm accounting and record keeping is a science of recording farm business transaction. It helps to understand nature and the extent of the financial effects. Records are important in analyzing farm business to understand the fairness of business and its weakness and improvements needed.

**Advantages of farm records and accounts**

- Means to higher income
- Basis for diagnosis and planning
- Way to improve managerial skills of farmer
- Basis for credit acquisition and management
- Guide to better home management
- Basis for conducting research in agricultural and production economics
- Basis for government policies

General forms of farm record are:

1. Farm inventory
2. Farm cash accounts or farm financial accounts
3. Classified farm cash account and annual farm business analysis
4. Supplementary financial records (Capital assets sale register; cash sale register; credit sale/purchase register; wage register; funds borrowed, repayments register, farm expenses register, non-farm income record)

### Farm inventory {-}

It is a list of the whole physical property of a business along with their values at a specified date. It is the complete list of farmers' assets. It is the first step in farm accounting. For an illustration, see Table \ref{tab:farm-inventory-ex}.

```{r farm-inventory-ex}
farm_inventory_ex <- readr::read_csv("./../data/farm_inventory_example.csv", skip = 3)

farm_inventory_ex %>% 
  select(-1) %>% 
  knitr::kable(caption = "Farm inventory of Betaaleshwor agro farm on Baisakh 2076 and Chaitra 2076", booktabs = TRUE, 
               col.names = c("", "", "Item", "Quantity", "Value", "Quantity", "Value"), escape = FALSE, longtable = TRUE) %>% 
  kableExtra::add_header_above(header = c(" " = 3, "Beginning of 2076" = 2, "End of 2076" = 2)) %>% 
  kableExtra::kable_styling(font_size = 8) %>% 
  kableExtra::collapse_rows(columns = c(1, 2)) %>% 
  kableExtra::column_spec(column = 1:7, width = c("8em", rep("6em", 2), rep("5em", 4)))
```


```{r farm-inventory-summary}
farm_inventory_ex %>% 
  summarise_at(vars(contains("value")), function(x)sum(x, na.rm = TRUE)) %>% 
  pivot_longer(cols = 1:2, names_to = "record_date", values_to = "value") %>% 
  mutate(record_date = str_remove_all(record_date, "_value") %>% 
           str_replace_all("_", " ") %>% 
           str_to_sentence()) %>% 
  add_row(tibble_row(record_date = "Change", value = .$value[1]-.$value[2])) %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(booktabs = TRUE, caption = "Change in inventory") %>% 
  kableExtra::kable_styling(latex_options = "hold_position")

```


\clearpage

# Calculation of depreciation of farm assets

## Objectives {-}

- To know different methods of calculation of depreciation

Decline in the value of an asset or capital equipment due to wear and tear of the asset, time, damages. Whenever any machines or equipment performs useful work, its wear and tear is bound to occur. Its efficiency reduces and becomes uneconomical to be used further and needs replacement by another new one.

**Concepts of depreciation**

- Original cost of asset: Bought price or purchasing cost of asset.
- Salvage value: The expected recovery of sales value of the asset at the end of useful life. (purchase value of asset/ expected life of that asset)
- Useful life: The expected time period for which the asset is to provide economic service i.e. the period for which the asset could be used for production.
- Depreciable cost: The original cost minus salvage value.
- Written down value: The value of asset at any point of time is the original cost less accumulated depreciation to date.

**Methods of computing depreciation**

1. Annual re-evaluation
2. Straight line method
3. Diminishing balance method
4. Sum of the year-digit method or reducing fraction method
5. Compound interest method

## Numerical {-}

1. (Using diminishing balance method) Calculate anuual depreciation for a tractor costing Rs. 240000 at the rate of 20\%.

*Solution*

We go on taking 20\% of the remaining balance until salvage value is reached, as in this case this is about Rs. 40330.

```{r}
annual_dep_fun <- function(initial_value = 24000, rate = 0.2, years = 7){
  years <- years
  annual_values <- vector("numeric", length = years)
  annual_values[1] <- initial_value
  for(y in 2:years){
    annual_values[y] <- annual_values[y-1] - annual_values[y-1]*0.2
  }
  return(annual_values)
}

tibble(
  "year" = 1:9, 
  "value_at_the_beginning_of_year" = NA, 
  "annual_depreciation" = NA, 
  "remaining_balance" = NA
) %>% 
  mutate(value_at_the_beginning_of_year = annual_dep_fun(initial_value = 240000, rate = 0.2, years = 9), 
         annual_depreciation = value_at_the_beginning_of_year-lead(value_at_the_beginning_of_year), 
         remaining_balance = value_at_the_beginning_of_year-annual_depreciation) %>% 
  rename_all(function(x)str_replace_all(x, "_", " ") %>% 
               str_to_sentence()) %>% 
  knitr::kable(booktabs = TRUE, caption = "Computation of annual depreciation by Diminishing balance method") %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = c("striped", "hold_position"))

```

<!-- 2. (Using compound interest method) The farmer may come to a problem of choice whether to buy new or old machine. For example, say new tractor cost Rs 5,00,000.00 and can run for 10 years (economic life). If he wants to buy old tractor now, it costs Rs 3,00,000.00 but can run for 5 years. After 5 years one more old tractor can be purchased in Rs 3,00,000.00 and can be operated for 5 years. Here for 10 years of operation, which one should he buy ? New one or two old ones ? -->

<!-- For this, discounting technique will help the farmer. Discount the future value into present value and compare. Let the interest rate be 10\% per annum. -->

<!-- ```{r} -->
<!-- interest_rate <- 0.1 -->
<!-- machine_life <- 5 -->
<!-- tribble( -->
<!--   ~"year", ~"value_of_new_machine", ~"value_of_first_old_machine", ~"value_of_second_old_machine", ~"present_value_of_two_old_machines",  -->
<!--   0, 500000, 300000, NA, NA,  -->
<!--   5, NA, NA, 300000, NA  -->
<!-- ) %>%  -->
<!--   mutate(present_value_of_two_old_machines = case_when( -->
<!--     year == 0 ~ value_of_first_old_machine,  -->
<!--     year == 5 ~ value_of_second_old_machine/(1 + interest_rate)^machine_life -->
<!--   )) %>%  -->
<!--   rename_all(function(x)str_to_sentence(str_replace_all(x, "_", " "))) %>%  -->
<!--   knitr::kable(booktabs = TRUE) %>%  -->
<!--   kableExtra::column_spec(column = 1:5, width = c("3em", rep("8em", times = 4))) -->
<!-- ``` -->

<!-- Although the value of two old tractors is Rs. 6,00,000 which is one lakh more than the new one, the analysis suggests us that the farmer should buy the two old ones because the discounted value of two old tractors will be less (Rs. 4,86,276) than the new one (Rs. 5,00,000). -->

<!-- $PV = \frac{P}{(1 + i)^n}$, present value of future sum, current value of sum of money to be received in future. -->

\clearpage

# Preparation of Balance sheet of a farm

## Objectives {-}

- To prepare balance sheet of a farm

## Introduction {-}

The balance sheet, also called net worth or financial statement, is a summary of assets and liabilities of a business together with the statement of the owners' equity or net worth. This is a listing of all assets and liabilities of a farm at certain point of time. This reports the financial strength and progress of the business. The term balance here implies that the value of the assets must be equal to the value fo liabilities plus owner's equity or net worth.

\begin{minipage}[c]{\textwidth}
\scalebox{0.8}{\noindent\begin{minipage}{0.3\textwidth}
\raggedleft
\begin{itemize}
\item Net worth or equity or net assets or capital of a company/farm is calculated by calculating the difference between the Total Assets and the Total Liabilities.
\item An example Net worth statement of MudnBrick Commercial Vegetable and Livestock Farm is presented in Table \ref{tab:net-worth-state}.
\end{itemize}
\end{minipage}}
\scalebox{1.0}{\noindent\begin{minipage}{0.7\textwidth}

```{r net-worth-state}
net_worth <- readxl::read_xlsx("../data/concepts_of_economics.xlsx", sheet = "net_worth_state", col_names = FALSE)
net_worth[-c(1,2,3,4),] %>% 
  knitr::kable(format = "latex",
             caption = "Net worth statement",
             col.names = c("", "", "", ""), booktabs = TRUE) %>% 
  kableExtra::kable_styling(position = "center", full_width = FALSE, latex_options = c("striped", "HOLD_position"), font_size = 8) %>% 
  kableExtra::add_header_above(c("Net Worth Statement" = 4), bold = T) %>%
  kableExtra::add_header_above(c("Financial condition as of 2011-11-11" = 4), italic = T) %>%
  kableExtra::add_header_above(c("MudnBrick Commercial Vegetable and Livestock Farm" = 4), bold = T) %>%
  kableExtra::row_spec(row = c(1,2,8,9,15,16,21,23,24), bold = TRUE, align = "c")
```

\end{minipage}}
\end{minipage}

\clearpage

# Preparation of Income statement of a farm

## Objectives {-}

- To be able to prepare an Income statement of a farm

## Introduction {-}

Balance sheet shows only the amount of wealth held by a business at one moment in time. Profit generated during a period is a main concern of many users of financial statements. P\&L statement provides a picture of profitability of the farm -- How much wealth (that is, profit) was generated, or lost, by the business over that period? 

Profit (loss) is defined as the increase (decrease) in wealth arising from trading activities. This is more accurately done using an accrual statement, which further accounts for inventory, accounts payable, and receivables, as well as depreciation expense. The difference between the total revenue and total expenses will represent either profit (if revenue exceeds expenses) or loss (if expenses exceed revenue). The period over which profit or loss is normally measured is usually known as the "accounting period", or "financial period".

```{r cost-return-an, echo=FALSE}
inc_state <- readxl::read_xlsx("../data/concepts_of_economics.xlsx", sheet = "cost_ret_analysis4", skip = 3, col_names = T)
inc_state %>% 
  mutate_at(.vars = vars(contains("amount")), function(x) kableExtra::cell_spec(x, "latex", bold = T, color = kableExtra::spec_color(x, end = 0.9))) %>% 
  knitr::kable(format = "latex",
             caption = "Income statement",
             booktabs = TRUE, escape = FALSE,
             col.names = c("", "Particulars", "Amount (Rs./ropani/annum)"), 
             align = "l") %>% 
  kableExtra::kable_styling(position = "c", full_width = FALSE, 
                            latex_options = c("striped", "HOLD_position"), font_size = 8) %>% # changed full width
  # for rowspec columns with na will be formatted first
  kableExtra::row_spec(row = which(is.na(inc_state[,1])), bold = TRUE, align = "c") %>%
  kableExtra::row_spec(row = 0, italic = TRUE, bold = TRUE) %>%
  kableExtra::add_header_above(c("Income statement" = 3), bold = T)
```


\clearpage

# Calculation of farm efficiency measures

## Objectives {-}

- Analyzing both physical and financial efficiency measures

## Introduction {-}

*Efficiency ratios*

1. Land use efficiency

$$\text{Production efficiency} = \frac{\text{Yield of the crop in a farm}}{\text{Average yield of the locality}}$$

$$\text{Cropping intensity} = \frac{\text{Area cropped}}{\text{Total cultivated area}}$$

$$\text{System index} = \frac{\text{Potential net income per hectare on farm}}{\text{Average standard net income per hectare in the area}} \times 100$$

2. Labor efficiency measures

- Crop acres per man or per man-year
- Livestock maintained per man or per man-year
- Gross profit per man or per man-year

3. Machinery efficiency

- Machinery and equipment cost per cropped acre
- Investment in machinery and equipment per crop acre

*Cost ratios*

1. Operating ratio: Represents the proportion absorbed by operating expenses out of the gross income.

$$\text{Operating cost ratio} = \frac{\text{Total operating cost or farm expenses}}{\text{Total profit or gross farm income}}$$

2. Fixed ratio

$$\text{Fixed ratio (aka. Overhead charges ratio)} = \frac{\text{Total fixed cost per year}}{\text{Gross income}}$$

3. Gross ratio

$$\text{Gross ratio} = \frac{\text{Total farm expenses}}{\text{Total cash expenses}}$$

4. Expense structure ratio

$$\text{Expense structure ratio} = \frac{\text{Fixed cash expenses}}{\text{Total cash expenses}}$$

*Profitability ratios*

1. Capital turnover ratio
2. Rate of return on debit and equity capital
3. Rate of return on equity capital

\clearpage

# Representation of production chain, market chain and supply chain in a value chain map of an agricultural commodity

## Objectives {-}

- To represent graphically the production chain, market chain and supply chain in a value chain map of an agricultural commodity

## Introduction {-}

Porter (1985) defined value chain is a chain of activities. Products pass through all activities of the chain in order and at each activity the product gains some value. The value chain categorizes the generic value adding activities of an organization.

A value chain map graphically illustrates all of the components, and relationships between them. Value chain maps demonstrate how a product in an industry moves from raw material through production, processing, and other steps, until it eventually ends with the consumer.

## Value chain map of ginger produced in Sindhuli district {-}

```{r diagram-value-chain-ginger, fig.width=4, out.width="60%", fig.pos="H", fig.align='center'}

DiagrammeR::grViz(
  "
  digraph boxes_and_circles {
  graph [overlap = true, fontsize = 16, layout=dot, rankdir=BT, compound=true]
  
  node [shape = rectangle, fontname = Helvetica, style = filled, color = DodgerBlue, fillcolor = Wheat, fixedsize = false] //sets as rectangle
  # if fixedwith=true, then width may be set to any value 1, 2, 3...
  edge[color = MidnigtBlue, arrowhead = diamond, arrowtail = none]
  
  subgraph cluster_inputsupply {
    label='Input supply';
    rank=same;
    node[shape = box, color = Blue, fillcolor = Cyan]
      '@@1-1'
      '@@1-2'
  }
  
  subgraph cluster_production {
    label='Production';
    rank=same;
    node[shape = box, color = Blue, fillcolor = Cyan]
      '@@2'
  }
  
  subgraph cluster_collection {
    label='Collection';
    rank=same;
    node[shape = box, color = Blue, fillcolor = Cyan]
      '@@3'
  }
  
  subgraph cluster_processing {
    label='Processing';
    rank=same;
    node[shape = box, color = Blue, fillcolor = Cyan]
      '@@4-1'
      '@@4-2'
  }
  
  subgraph cluster_trading {
    label='Trading';
    rank=same;
    node[shape = box, color = Blue, fillcolor = Cyan]
      '@@5-1'
      '@@5-2'
      '@@6'
  }
  
  subgraph cluster_consumption {
    label='Consumption';
    rank=same;
    node[shape = box, color = Blue, fillcolor = Cyan]
      '@@7-1'
      '@@7-2'
  }
  
  '@@1-1' -> '@@2' [lhead = cluster_inputsupply, label='Price of inputs', constraint=true]
  '@@2' -> '@@3' [label = 'Farm gate price (Rs 60/kg)']
  '@@3' -> '@@4-1' [label = 'Collector\\'s price (+ Rs 5/kg)']
  '@@3' -> '@@4-2'
  '@@4-1' -> '@@5-1' [label = 'Processor\\'s price (+ Rs 5-20/kg)']
  '@@4-1' -> '@@5-2'
  '@@4-2' -> '@@5-1'
  '@@4-2' -> '@@5-2'
  '@@4-2' -> '@@6'
  '@@5-1' -> '@@6'
  '@@5-1' -> '@@6'
  '@@5-2' -> '@@7-1'
  '@@5-2' -> '@@7-2'
  '@@6' -> '@@7-1' [label = 'Trader\\'s price (+ Rs 5-10/kg)']
  '@@6' -> '@@7-2'
  }
  
  [1]: c('Agriculture knowledge center, cooperatives', 'Private sector input suppliers')
  [2]: 'Ginger growers'
  [3]: 'Cooperatives, wholesaler, traders'
  [4]: c('Pickles, candy, ginger powder and other processed ginger products', 'Fresh ginger')
  [5]: c('District wholesaler traders', 'Export to india')
  [6]: 'Retailer'
  [7]: c('Consmption of fresh ginger', 'Consumption of processed ginger products')
  "
)
```

\clearpage

# Determination of project worth using investment appraisal criteria

## Objectives {-}

- To determine projects' worth using investment appraisal criteria
