---
author: Deependra Dhakal
fontsize: 12pt
title: "Agribusiness environment and management system"
date: \today
institute:
  - GAASC, Baitadi
  - Tribhuwan University 
output: 
  binb::iqss:
    keep_tex: true
    slide_level: 2
    includes:
      in_header: iqss_beamer_header.tex
classoption: "aspectratio=169"
---

```{r setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2, knitr.kable.NAN = "")
options(kableExtra.latex.load_packages = FALSE)
```

# Meaning and definition

## Environment

- It is the situation in which agribusiness firm is operated.
- It checks the potentiality and feasibility of business for overall investment
- One of the important steps in business initiation
- Analysis is conducted throughout the business period
- The environmental factors or forces which affect the success of agribusiness are:
  1. Economic environment
  2. Demographic environment
  3. Socio-cultural environment
  4. Technological environment
  5. Political environment
  6. Legal environment
- Overall, it is resultant of those components like government, technology, management, organization and individuals (public).

##

- In agribusiness there are several producers and several consumers as well. So the effect of single producer and consumer is negligible; i.e. they have no single effect on fixing market price, demand and supply.
- Therefore the market equilibrium does not change unless external forces are operating.

## Forces affecing agribusiness

- technology changes,
- production (yields per acre, rate of gain per animal, pounds of milk per animal),
- local labor (availability and price),
- land and machinery (availability and price),
- consumer and industry demand for your products (tase and preference),
- markets (availability, niche, traditional, organic),
- operating environmental forces,
- conservation programs (regulations),
- family living and medical expenses (current and future),
- macroeconomic forces (inflation, interest, and exchange rates),
- trade agreements (GATT, SAFTA, BIMSTEC), and
- other items unique to the farm business. 

## Importance of agribusiness environment analysis

- The manager needs to be dynamic to effectively deal with the challenges of the environment as the business environment is non-static. Some of the benefits of environment
scanning are:
1. It creates an increased general awareness of environment changes on the part of management.
2. It guides with greater effectiveness in matters relating to government.
3. It helps in marketing analysis.
4. It suggests improvement in diversification and resources allocation
5. It helps firms to identify and capitalize upon opportunities rather than losing out to competitors.
6. It provides base of objective qualitative information about the business environment that can subsequently be of value in designing the strategies.

# Management system and managerial decisions

## Agribusiness management functions

\begin{columns}

\column{0.6\textwidth}

\footnotesize
1. Planning

\begin{itemize}
\item What to produce ? How to produce ? How much to produce ?
\item Set objectives, formulate product production, processing or marketing strategies and plan accordingly
\item Prepare work schedule and budget accordingly
\item Plan joint venture
\item First prepare managerial work plan and then work under it.
\end{itemize}

\column{0.4\textwidth}

```{r management-functions, fig.width=2, fig.height=3, out.width="28%"}
DiagrammeR::mermaid(
  "
  graph TB
  
  A((Directing)) --> B((Controlling))
  B --> C((Planning))
  C --> D((Organizing))
  D --> E((Staffing))
  E --> A
  "
, width = 400, height = 400)
```


\end{columns}

##

2. Organizing

- Systematic classification and grouping of human and other resources in a manner consistent with the firm's goals.
- Organizing task occurs continuously throughout the life of a firm
- Defining line of authority and responsibility
- Decentralization of authority
- Coach and train all employees

##

3. Staffing

- Process of recruiting, selecting, training and developing human resources in organization

4. Directing 

- Process of influencing employees to work for the achievement of the organizational goals

5. Controlling

- The collection of relevant feedback information, analyzing of such data and as per need taking of corrective action
- It measures current performance of the firm, cooperatives or company.

## Best management practices

- The best management practices are research driven, industry based and proven in practice.
- Enlisting:
  1. Visioning
  2. Risk management
  3. Human resource management
  4. Technically innovative business
  5. Market proactive and consumer oriented (marketing management)
  6. Addressing sustainable development practices

## Decision-making at the level of top management

- Determine the internal allocation of most of the resources. 
- Two criteria are set: 
  - The first is a budgetary (financial) criterion, i.e. are there fund available for the realization of the proposed agribusiness project? 
  - The second is an improvement criterion, i.e. does the agribusiness project being proposed improve the existing situation beyond doubt?

## Decision-making at lower levels of management (Administration)

- Adaptive rational system, i.e. on the floor, day-to-day activities, 'blue print' and rule-of-thumb.
- It involves choosing the best course of action (choosing among various sets of alternatives) for accomplishing goals out of available alternatives.
