---
author: Deependra Dhakal
fontsize: 12pt
title: "Cooperative operation in commercial farming"
date: \today
institute:
  - GAASC, Baitadi
  - Tribhuwan University 
output: 
  binb::iqss:
    keep_tex: true
    slide_level: 1
    includes:
      in_header: iqss_beamer_header.tex
classoption: "aspectratio=169"
---

```{r setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2, knitr.kable.NAN = "")
options(kableExtra.latex.load_packages = FALSE)
```

# Commercial farming

- Cultivation of crops, rearing of livestock and producing farm commodities for purpose of generating profit.
- Commercial farming makes use of cost reducing technologies in production and optimize production levels based on consumer demand and price of produce.
- These types of agricultural systems are more specialized in choice of crop or enterprise, so that operational plan may learn from industry sector.
- Term generally refers to high input use condition where inputs such as irrigation, fertilizer and land management are heavily controlled to aim for higher volume of output.
- Cooperative farming ensures better production of agricultural goods through encouragement of use of bigger scale technology, besides also providing chance for capital formation among smallholder farmers.

# Cooperative: Lesson from India

\footnotesize

- Dairy farming based on the Anand Pattern, with a single marketing cooperative, is India's largest self-sustaining industry and its largest rural employment provider.
- Successful implementation of the Anand model has made India the world's largest milk producer.
- Here small, marginal farmers with a couple or so heads of milch cattle queue up twice daily to pour milk from their small containers into the village union collection points. 
- The milk after processing at the district unions is then marketed by the state cooperative federation nationally under the Amul brand name, India's largest food brand. 
- With the Anand pattern three-fourth of the price paid by the mainly urban consumers goes into the hands of millions of small dairy farmers, who are the owners of the brand and the cooperative. 
- The cooperative hires professionals for their expertise and skills and uses hi-tech research labs and modern processing plants and transport cold-chains, to ensure quality of their produce and value-add to the milk. 

# Cooperative farming

\footnotesize

```{r mermaid-cooperative-farming, fig.width=6, out.width="80%", fig.align='center'}
DiagrammeR::mermaid("
                    graph LR
                    
                    A(Types of cooperative farming)--> B(Cooperative better farming)
                    A --> C(Cooperative tenant farming)
                    A --> D(Cooperative collective farming)
                    A --> E(Cooperative joint farming)
                    ")
```

#

1. Cooperative joint farming society

- Most comprehensive type of cooperative farming society.
- Small owners pool their land for the purpose of joint cultivation.
- The ownership is individual but the operations are collective.
- The management is democratic and is elected by the members of the society.
- It purchases various inputs from the market and arranges for the marketing of the produce.
- It also seeks financial assistance from outside agencies to carry on these activities.

#

2. Cooperative better farming society
- The members do not cultivate their land jointly. Each member cultivates his own land.
- However, they co-operate with each other for pre-sowing and post harvesting operation.
- For instance, they purchase various agricultural inputs like seeds, fertilizers, insecticides, services of machinery etc. on cooperative basis.
- They sell the crops jointly.
- A cooperative better farming society may also arrange for financial assistance for carrying on these activities.
- The members pay for the service rendered to them by the society.

# 

3. Cooperative tenant farming society
- This is a society which purchases or leases in land from the Government or private individuals and then in turn leases out the land to its members.
- Such societies are usually organized by landless farmers.
- The members cultivate the land and pay the rent falling to their share, to the society.
- The society also renders various other services to its members and charges from, its members for the service so rendered.
- The profits earned by the society are distributed among its members according to some agreed formula.

#

4. Cooperative collective farming society
- This type of society involves pooling of their land by the members on a permanent basis.
- A member who joins this society cannot ever withdraw his land from the society.
- He can only transfer his land to some other person who will now become a substitute member of the society.
-The functions of this society are similar to those performed by a
cooperative joint farming society.
- The member get their wages and profits according to the labor and land respectively contributed by them.
- It is obvious that such a society is formed in contravention of the general principles of cooperative i.e. voluntary membership with a right to withdraw from the society at any time.

# Advantages of cooperatives in commercial farming

\footnotesize

1. **Increases bargaining strength of the farmers**: Many of the defects of the present agricultural marketing system arise because often one ignorant and illiterate
farmer (as an individual) has to face well-organized mass of clever intermediaries. If the farmers join hands and for a co-operative, naturally they will be less prone to exploitation and malpractices. Instead of marketing their produce separately, they will market it together through one agency.

2. **Direct dealing with final buyers**: The co-operatives can altogether skip the intermediaries and enter into direct relations with the final buyers. This practice will eliminate exploiters and ensure fair prices to both the producers and the consumers.

3. **Provision of credit**: The marketing co-operative societies provide credit to the farmers to save them from the necessity of selling their produce immediately after harvesting. This ensures better returns to the farmers.

#

\footnotesize

4. **Easier and cheaper transport**: Bulk transport of agricultural produce by the societies is often easier and cheaper. Sometimes the societies
have their own means of transport.
5. **Storage facilities**: The co-operative marketing societies generally have storage facilities. Thus the farmers can wait for better prices.
6. **Grading and standardization**: This task can be done more easily for a co-operative agency than for an individual farmer. For this purpose, they can seek assistance from the government or can even evolve their own grading arrangements.
7. **Market intelligence**: The co-operatives can arrange to obtain data on market prices, demand and supply and other related information from the markets on a regular basis and can plan their activities accordingly.

#

8. **Influencing marketing prices**: Wherever strong marketing co-operative are operative, they have bargained for and have achieved, better prices for their agricultural produce.
9. **Provision of inputs and consumer goods**: The co-operative marketing societies can easily arrange for bulk purchase of agricultural inputs, like seeds, manures fertilizers etc. and consumer goods at relatively lower price and can then distribute them to the members.
10. **Processing of agricultural produce**: The co-operative societies can undertake processing activities like crushing seeds, ginning and pressing of cotton, etc. In addition to all these advantages, the co-operative marketing system can arouse the spirit of self-confidence and collective action in the farmers without which the programs of agricultural development, howsoever well-conceived and implemented, holds no promise to success.

# Registration of cooperatives in Nepal

\footnotesize

A cooperative organization can carry out its functions only after its registration under the Cooperative act, 2074. For registration of cooperatives following procedures should be followed.

1. Preliminary meeting: According to the Cooperative act, 2074, there should be at least 25 members to form a cooperative society. Preliminary meeting must be held before applying for the registration. The meeting is held in the presence of 25 members under 1 chairman among them. The following things should be discussed in the meeting:

- Commencement of the business
- The name and address of the society
- The objectives of the society
- The value of each share
- Membership fee

#

\footnotesize

2. Filing an application for registration: After preparing and passing proposed by laws and working schemes in the preliminary general meeting. In application should be submitted to the office of registrar, department of cooperatives, and government of Nepal

Following things are mentioned in the application form:

- Proposed name of society
- Address
- Objectives
- Working areas
- Liabilities
- Total share capital
- Total number of shares to be paid
- Two copies of law of proposed society
- Original copy of working scheme
- Copies of citizen certificate
- Application must be signed by chairman

#

3. Receiving the certificate of registration

After filing application for registration, certificate of registration is to be received. After applying application along with document, they are submitted at the registration office. Then the registrar checks all the documents. If the documents are satisfactory then registrar will issue certificate of registration. After receiving certificate of registration, the society can operate.

# Legal provisions for dissolution of co-operative society

Cooperative society can be dissolved under following circumstances:

- Two third majority of total number of society can take decisions of dissolution.
- Registrar can dissolve it, if application with reasonable clause is received.
- Registrar can dissolve it, if the society is found inactive and not operating for two years.
- The registrar can dissolve it, if it is found operating against the law and objectives of it.
