---
author: Deependra Dhakal
fontsize: 12pt
title: "Cooperative: History and development in Nepal"
date: \today
institute:
  - GAASC, Baitadi
  - Tribhuwan University 
output: 
  binb::iqss:
    keep_tex: true
    slide_level: 2
    includes:
      in_header: iqss_beamer_header.tex
classoption: "aspectratio=169"
---

```{r setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2, knitr.kable.NAN = "")
options(kableExtra.latex.load_packages = FALSE)
```

# History

##

- The prehistoric form of cooperatives in Nepal were Kipat, Guthi, Parma, Dhikuri, Dharma bhakari, etc. The characteristics of informal institutions were similar to cooperatives.
- The GON established the Department of Cooperatives under the Ministry of Planning, Development, and Agriculture in 2010 BS
- The department started building for cooperative movement
- The modern cooperative started from Rapti valley of Chitwan district as a part and partial flood relief and resettlement program.
- First cooperative society established in Shardanagar, Chitwan -- "Bakhan sahakari sanstha"

##

- The cooperatives act was revised several times.
- In 2016, the cooperative organization act and in 2018 B.S. the cooperative organization regulations were promulgated.
- In 2018, Sajha Publications, Sajha Travels, Sajha Health Service were formed under the Central Cooperative Sajha Organization.
- After the restoration of democracy, Sajha Organization Act 2041 was dissolved and Cooperative Act 2048 and Cooperative Regulations 2048 were passed by the parliament. These two laws are still in effect.
- The department of cooperatives has provided the authority for registration and regulation of cooperatives societies/unions/federations under the act.
- The department of cooperatives after establishment faced unification, merge, separation problem from aligned ministries.
- It is under Ministry of Land management, Agriculture and cooperatives, 2074/75. Before it was under Ministry of Cooperatives and Poverty Alleviation, 2069

##

- The co-operative act, 1992 
  - provided an opportunity for the Nepalese people to establish independent and autonomous co-operative societies by themselves.
  - has foundation on basic cooperative principles and has also provided a legal base both for establishment of cooperative societies and unions and federations.
- It is estimated that some three million people are affiliated in more than thirty thousand cooperatives, and more than 54 thousand people are employed directly in Cooperative business in Nepal.
- Agriculture  related cooperatives were the second most (25.88%) prevalent nature of cooperatives after the saving and credit cooperatives (MoCPA, 2015).

# Cooperative: Developments in Agriculture

##

- Nepal's new constitution (Constitution of Nepal, 2070) makes strong commitments to cooperatives, many of which are agricultural.
- It has been spoken often that cooperative in Nepal can provide an effective institutional framework to promote and sustain development at local level.
- Also, cooperative in Nepal hold promise to break the vicious cycle of rural poverty through contributing to economic and social improvements.
- Current development policies and plans of Nepal have considered the cooperative sector as one of the 3 pillars of the country's economic development; other two being the public sector and the private sector.
- Such policies and plans have adopted a strategy of encouraging small and marginal farmers to produce low volume, high value crops and  commodities  by  organizing  them into farmer  groups  and  co-operatives. 

##

- Agricultural cooperatives offer market opportunities to smallholders, and provide them with services such as better training in natural resource management and better access to information technologies, innovations and extension services, thus helping smallholders achieve sustainable livelihoods and improve food security in their communities (IFAD 2012).
- As in other developing countries, agricultural cooperatives in Nepal are involved in overall socio-economic well being of farming community, in general, and members, in particular. 

# Role of agricultural cooperatives

##

```{r role-of-agricultural-cooperatives, fig.width=6, out.width="85%"}
DiagrammeR::mermaid(
"
graph LR
subgraph 
A[Promoting commercialization of agriculture through<br> provision of farm inputs, credit and improved technologies to members]
B[Small scale processing and value addition of primary agricultural products]
C[Marketing of primary as well as processed agricultural products<br> in domestic as well as international markets]
D[Capacity building of member farmers in technical<br> as well as management aspects of agricultural production and marketing]
E[Promotion of sustainable use of local resources]
F[Community development]
G[Policy lobbying for farmer friendly agriculture]
H[Networking with government and non-government service providers<br> to tap resources for benefit of farming communities]
I[Promotion of social inclusion and social harmony in the community]
end
"
)

```

##

- Poverty is considered a major threat for world peace and stability.
- Achievements made by the cooperatives in economic prosperity of societies, particularly in developing countries:
  - Dairy cooperatives in India and agricultural cooperatives in Zambia;
  - ICT cooperatives in rural America after the World War II, Poland, Albania, Argentina and Bolivia;  
  - Electric service cooperatives in the Philippines, Bangladesh India and Bolivia; 
  - Financial service cooperatives in Mexico, the  Philippines, Bolivia, Nicaragua, Romania, Poland, Ecuador, El Salvador, Guatemala, Honduras, Jamaica and Russia;  

##

- There has been a remarkable growth in the number of cooperatives in recent years and one of main reasons for such growth is attributed to the government extension and farm sector policy of supporting farmers through the cooperatives. 
- For example, government policy of distributing subsidized fertilizers through the agriculture and related cooperatives has spurred the trend of forming cooperatives in rural and peri-urban areas.
- Government budget for the fiscal year 2008/09 introduced a new program "Cooperatives in every village, food storage in every house" to increase the agricultural productivity through cooperatives(MOF, 2009). This program aimed at motivating small farmers to commercialize their agriculture through cooperative models.
- The government allotted Rs. 85million for providing incentives to cooperative farming.

##

- Budget for fiscal year 2011/12 has given more importance to the cooperatives; it has brought various cooperatives programs and policies for the promotion and expansion of cooperatives as one of the central pillars of the economy. 
- This budget launches a national program of "Cooperatives in Villages and Cites, Employment at Every Household" to encourage cooperative farming to increase agricultural productivity.
- The government proposed Rs. 10 million for this program.
- The budget provided credit facility at a concessional interest rate through small farmer cooperatives for the people involved in livestock, poultry, and fish farming. It also aimed to provide 50 percent subsidy on the capital expenditure to the small farmer cooperatives in the purchase of machinery and equipment (MOF, 2012).
- The government budget for Fiscal Year 2016/17 said, "Activities of agricultural production, process and marketing will be conducted in such areas with the involvement of agricultural entrepreneurs, cooperatives, the private sector, and the Government."

## Agricultural cooperative: A case study from Nepal

- For case studies of Agricultural Cooperatives in Nepal, refer to [Building social capital within the framework of agricultural cooperatives development in rural Nepal](https://core.ac.uk/download/pdf/43382733.pdf) by Ram Krishna Shrestha.
- For types of farmer cooperatives, refer to [Farmer Cooperatives for Food Self-sufficiency, Agricultural Commercialization, and the Socio-economic Development of Nepal](http://cffn.ca/old/conf/2007/Farmer-Cooperatives-for-Food-Self-Sufficiency-DDPoudel-cffn-canada.pdf)

## Future direction for cooperatives

- Cooperatives here must at least match the large proportion of agricultural output which is handled by cooperatives in Europe and North America, where they play a major role. If this happens, Nepal's agricultural cooperatives can indeed be an important force for development, contributing to the achievement of the Sustainable Development Goals (which, under goal 2, address farm incomes).
- Bulk-buying inputs (principally seed, fertilizer, and insecticide) keeps costs down since cooperatives can get better deals with suppliers than individual farmers can. 
- Selling output in bulk also means better prices, particularly if the cooperative can consistently supply buyers with the amounts they need. Market intermediaries often complain that small farmers cannot reliably do this and that quality is often poor. Cooperatives' ability to provide acceptable quality is therefore also very important.
